---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Saima Arina"
date: "02/04/2026"
output: github_document
---

### Load packages and data


```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r dimensions of Dennys dataset}
nrow(dennys)
ncol(dennys)
dim(dennys)
```
The Denny's dataset has 1,643 rows and 6 columns. Each row represents a Denny's location. The variables are: address, city, state, zip, longitude, and latitude. 

### Exercise 2

```{r dimensions of Laquinta dataset}
nrow(laquinta)
ncol(laquinta)
dim(laquinta)
```
The La Quinta dataset has 909 rows and 6 columns. Each row represents a La Quinta location. The variables are: address, city, state, zip, longitude, and latitude. 

### Exercise 3

There are La Quinta locations in Canada, Mexico, China, New Zealand, Georgia (the country), Turkiye, the United Arab Emirates, Colombia, and Ecuador. There are no Denny's locations outside of the United States. 

### Exercise 4

To find a location outside the US using just the data, you could filter for any location that either has no state or an abbreviated state that does not match any of the states of the USA. 

### Exercise 5

```{r Filter for states that are not in states$abbreviation}
dennys %>%
  filter(!(state %in% states$abbreviation))
```
There are no Denny's locations outside of the US. 

### Exercise 6

```{r Add a country variable to the Denny’s dataset and set all observations equal to "United States"}

dennys %>%
  mutate(country = "United States")

dennys <- dennys %>% 
  mutate(country = "United States")
```


 
### Exercise 7

There are La Quinta locations in Canada: British Columbia (Kelowna and Richmond), and Ontario (Oshawa). In Mexico: Aguascalientes, Ciudad Juarez, Poza Rica, Puebla, Reynosa, and San Jose Chiapa. In China: Chengdu, Qionghai, Suzhou, Taiyuan, Turpan, Weifang, and Zunyi. In New Zealand: Auckland and Queenstown. In Georgia: Batumi. In Turkiye: Bodrum, Cesme, Giresun, and Istanbul. In United Arab Emirates: Abu Dhabi and Dubai. In Colombia: Medellin. In Ecuador: Quito. 

